---
title: "FEEDBACK ORIENTATION STUDY: preliminary analysis"
author: "Stefanie Meliss"
date: "`r Sys.Date()`"
output: html_document
---

For the analysis below, exclusion criteria for participants in the feedback orientation study were applied.  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# libraries
library(kableExtra)
library(dplyr)
library(GGally)

# empty work space
rm(list = ls())

# get directories
dir <- getwd()

# process data: assumes this has been run
source("C:/Users/stefanie.meliss/OneDrive - Ambition Institute/code/colour_palette.R")
# source(file.path(dir_root, "1_prepare_pretest_data.R"))

# function to determine outliers
devtools::source_url("https://github.com/stefaniemeliss/fo_sim/blob/master/helper_functions.R?raw=TRUE")

# read in data
file = file.path(dir, "pretest_ARK.xlsx")
df <- xlsx::read.xlsx(file, sheetName = "SCALES")

# remove ppt
file  <- "C:/Users/stefanie.meliss/OneDrive - Ambition Institute/research_projects/Ark Feedback Orientation and Sims/Analysis/Final_exclusions_Ark_ITTs_6Sept.xlsx"
tmp <- xlsx::read.xlsx(file = file, sheetIndex = 2)
to_remove <- tmp$pseudonym

df <- df %>% filter(!pseudonym %in% to_remove)


```


### Questionnaire data

The sum score for each scale was computed following instructions given by the scale author(s). We have the same number of observations for each scale.

```{r scales, echo = F, results='asis'}
scales <- names(df)[grep("fo|conscient|m_a|p_a|task_|pref", names(df))]

# compute descriptives for fo subscales
tmp <- psych::describe(df[, scales])
# round values
tmp <- round(tmp, 2)
# add scale name
tmp$vars <- row.names(tmp)

# print to markdown
kbl(tmp, caption = "Scale descriptives", row.names = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```

To better understand how the data is distributed and whether scales correlate with each other, a scatterplot matrix with embedded histrograms is produced. The results suggests that FO does not correlate with any neighbouring constructs.  


```{r, echo=FALSE, results='asis', fig.align='center', fig.height=12, fig.width=12, message=F, warning=F}
# save data for matrix in seperate df
desc <- df[, c("pseudonym", scales)]

# define plotting function for lower diagonal
lowerfun <- function(data,mapping){
  ggplot(data = data, mapping = mapping)+
    geom_point(position=position_jitter(height=0.01, width=0.01), col = navy) 
  # +
  # coord_cartesian(ylim = c(-2,2), xlim = c(-2,2))
  
} 

# define plotting function for upper diagonale
diagfun <- function(data,mapping){
  
  ggplot(data = data, mapping = mapping)+
    geom_histogram(stat = "count", binwidth = 1, fill = coral) 
  # + 
  #   coord_cartesian(ylim = c(0,100), xlim = c(-2,2))
} 

# create plot
plt <- ggpairs(desc[, -1], lower = list(continuous = wrap(lowerfun)), diag = list(continuous =  wrap(diagfun))) + theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), strip.text.y.right = element_text(angle = 360), strip.text.x.top = element_text(angle = 90))



# plotly::ggplotly(plt)
# print plot
print(suppressWarnings(plt))

```

### n-back data

To process n-back data, trial-level data was used (after removing practice trails). For each participant, event counts (n-back trial vs. non n-back trial) and performance according to signal detection theory (i.e., hits, misses, false alarms) were extracted. Additionally, rates for each event/performance type were computed. Accuracy rate is computed as hit rate - false alarm rate.  
  

```{r nback, echo = F, results='asis'}
nback <- names(df)[grep("total_|rate_", names(df))]

# compute descriptives for fo subscales
tmp <- psych::describe(df[, nback])
# round values
tmp <- round(tmp, 2)
# add scale name
tmp$vars <- row.names(tmp)

# print to markdown
kbl(tmp, caption = "n-back descriptives", row.names = F) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```

```{r, echo=FALSE, results='asis', fig.align='center', fig.height=12, fig.width=12, message=F, warning=F}

# remove outlier
desc <- df[, c("pseudonym", nback)]

# redo scatterplot matrix
# create plot
plt <- ggpairs(desc[, -1], lower = list(continuous = wrap(lowerfun)), diag = list(continuous =  wrap(diagfun))) + theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), strip.text.y.right = element_text(angle = 360), strip.text.x.top = element_text(angle = 90))

print(suppressWarnings(plt))

```